---
title: "A Reanalysis of Livestock Forage Disaster Program Eligibility, 2012--2024"
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

## Introduction

```{r}
#| label: setup
#| echo: false

# update.packages(repos = "https://cran.rstudio.com/",
#                 ask = FALSE)

# install.packages("pak",
#                  repos = "https://cran.rstudio.com/")

# installed.packages() |>
#   rownames() |>
#   pak::pkg_install(upgrade = TRUE,
#                  ask = FALSE)

# pak::pak(
#   c(
#     "arrow?source",
#     "sf?source",
#     "curl",
#     "tidyverse",
#     "tigris",
#     "rmapshaper",
#     "furrr",
#     "future.mirai",
#     "dtplyr",
#     "ggspatial"
#   )
# )

library(magrittr)
library(tidyverse)
library(sf)
library(arrow)
library(furrr)
library(future.mirai)
library(dtplyr)

sf::sf_use_s2(TRUE)

dir.create("data-raw",
           showWarnings = FALSE,
           recursive = TRUE)

dir.create("data-derived",
           showWarnings = FALSE,
           recursive = TRUE)

dir.create(
  file.path("data-raw", "counties", "census"),
  showWarnings = FALSE,
  recursive = TRUE
)

dir.create(
  file.path("data-derived", "counties"),
  showWarnings = FALSE,
  recursive = TRUE
)

```

```{r}
#| label: get-usdm-dates

get_usdm_dates <-
  function(as_of = lubridate::today()){
    as_of %<>%
      lubridate::as_date()
    
    usdm_dates <-
      seq(lubridate::as_date("20000104"),
          lubridate::today(),
          "1 week")
    
    usdm_dates <- usdm_dates[(as_of - usdm_dates) >= 2]
    
    return(usdm_dates)
  }

usdm_dates <-
  get_usdm_dates()


# Generate counties from the Normal Grazing Periods dataset
counties <-
  readr::read_csv(
    "https://data.sustainable-fsa.com/fsa-normal-grazing-period/fsa-normal-grazing-period.csv"
  ) |>
  dplyr::transmute(
    FIPS = stringr::str_c(`FIPS State Code`, `FIPS County Code`),
    State = `FIPS State Name`,
    County = `FIPS County Name`) |>
  dplyr::arrange(FIPS) %>%
  dplyr::mutate(County = ifelse(County == "Cata\xf1o", NA, County)) %>%
  dplyr::filter(!is.na(County)) %>%
  dplyr::distinct()


arrow::read_parquet("data/lfp.parquet") %>%
  dplyr::group_by(FIPS,
                  source,
                  `Program Year`,
                  `Pasture Type`) %>%
  dplyr::filter(Payments == max(Payments)) %>%
  # dplyr::group_by(FIPS,
  #                 `Program Year`,
  #                 `Pasture Type`) %>%
  # dplyr::filter(!all(Payments == Payments[1])) %>%
  dplyr::ungroup() %>%
  dplyr::select(!c(`Qualifying Drought Event`, `Qualifying Date`)) %>%
  tidyr::pivot_wider(names_from = source,
                     values_from = Payments) %>%
  dplyr::left_join(counties)


```

### Summary Recommendations

#### For FSA

- Review agency interpretation of eligible payments. Ensure regulation regarding capping payments by length of normal grazing period conforms with regulatory and legislative language, and is well documented.
- Use authoritative county boundaries from the US Census, updated annually, and ensuring consistency during the LFP program year.
- Assess the current FSA interpretation of "any part of a county" regulatory and legislative language. Ensure that procedures align with agency interpretation and are documented.
- 

#### For USDM

- Do not clip USDM polygons to boundaries, or do so with a buffer outside the land area.
- Ensure USDM data and metadata are archived in FGDC-compliant formats in a Federal repository.
- Ensure USDM data are spatially valid, and refine data preparation protocols to ensure validity.

#### For County Aggregation




## Data

### FSA Normal Grazing Periods

```{r}
fsa_normal_grazing_period <-
  readr::read_csv(
    "https://data.sustainable-fsa.com/fsa-normal-grazing-period/fsa-normal-grazing-period.csv"
  ) |>
  dplyr::transmute(
    `Program Year` = as.integer(`Program Year`), 
    FIPS = stringr::str_c(`FIPS State Code`, `FIPS County Code`),
    `Pasture Type`,
    `Grazing Period Start Date`,
    `Grazing Period End Date`,
    `Normal Grazing Period` = 
      lubridate::interval(
        start = `Grazing Period Start Date`,
        end = `Grazing Period End Date`
      )) |>
  dplyr::arrange(FIPS, `Program Year`, `Pasture Type`)
```


### USDM County Aggregations

```{r}
#| label: usdm-county-aggregations

usdm_county_aggregations <-
  list(
    usdm_reported = 
      file.path("https://data.sustainable-fsa.com/usdm-counties-reported",
                "usdm-counties-reported.parquet"),
    fsa_lfp = 
      file.path("https://data.sustainable-fsa.com/usdm-counties-fsa-lfp",
                "usdm-counties-fsa-lfp.parquet"),
    census = 
      file.path("https://data.sustainable-fsa.com/usdm-counties",
                "usdm-counties.parquet")
  ) |>
  purrr::map(arrow::read_parquet)

```


```{r}
#| label: process-usdm-county-max-class

process_usdm_county_max_class <-
  function(x){
    x %>%
      dtplyr::lazy_dt() %>%
      dplyr::group_by(
        FIPS = paste0(STATEFP, COUNTYFP),
        date = usdm_date
      ) %>%
      dplyr::summarise(
        usdm_class = max(usdm_class),
        .groups = "drop"
      ) %>%
      dplyr::arrange(
        FIPS, 
        date
      ) %>%
      dplyr::collect()
  }

usdm_county_max_class <-
  usdm_county_aggregations %>%
  purrr::map(process_usdm_county_max_class)


good_fips <-
  usdm_county_aggregations$usdm_reported %>%
  dplyr::select(STATEFP, COUNTYFP, State, County) %>%
  dplyr::distinct()

test <-
  usdm_county_max_class %>%
  dplyr::bind_rows(.id = "source") %>%
  dtplyr::lazy_dt() %>%
  dplyr::group_by(FIPS, date) %>%
  dplyr::filter(length(unique(usdm_class)) != 1L) %>%
  dplyr::collect() %>%
  tidyr::pivot_wider(names_from = source,
                     values_from = usdm_class) %>%
  na.omit()

test %>%
  dplyr::filter(date >= "2012-01-01",
                usdm_reported >= "D2" | fsa_lfp >= "D2") %>%
  print(n = 100)

usdm_county_aggregations$usdm_reported %>%
  dplyr::filter(STATEFP == "05",
                COUNTYFP == "039",
                usdm_date == "2015-10-06")

usdm_county_aggregations$fsa_lfp %>%
  dplyr::filter(STATEFP == "05",
                COUNTYFP == "039",
                usdm_date == "2015-10-06")

map_county_diff <-
  function(state = "05",
                county = "039",
                usdm_date = "2015-10-06"){
    paste0("https://data.sustainable-fsa.com/usdm/data/parquet/USDM_",usdm_date,".parquet") %>%
    sf::read_sf() %>%
  mapview::mapview(zcol = "usdm_class",
                   burst = TRUE) +
  mapview::mapview(
    sf::read_sf("https://data.sustainable-fsa.com/usdm-counties-fsa-lfp/data/fsa-lfp-counties.parquet") %>%
      dplyr::filter(STATEFP == state,
                    COUNTYFP == county)
  )
  }


map_county_diff(state = "01",
                county = "001",
                usdm_date = "2022-12-06")


```

#### USDM-reported Aggregations
```{r}
#| label: usdm-reported-county-aggregations

usdm_reported_county_aggregations <-
  file.path("https://data.sustainable-fsa.com/usdm-counties-reported",
            "usdm-counties-reported.parquet") |>
  arrow::read_parquet()

```

#### US Census Aggregations

```{r}
#| label: usdm-census-county-aggregations

usdm_census_county_aggregations <-
  file.path("https://data.sustainable-fsa.com/usdm-counties",
            "usdm-counties.parquet") |>
  arrow::read_parquet()

```


#### FSA LFP County Aggregations

```{r}
#| label: usdm-fsa-lfp-county-aggregations

usdm_fsa_lfp_county_aggregations <-
  file.path("https://data.sustainable-fsa.com/usdm-counties-fsa-lfp",
            "usdm-counties-fsa-lfp.parquet") %>%
  arrow::read_parquet()

```


### LFP Payment Files

```{r}

lfp_payments <-
  arrow::s3_bucket("sustainable-fsa/fsa-payment-files",
                   anonymous = TRUE,
                   region = "us-west-2") %>%
  arrow::open_dataset() |>
  dplyr::filter(`Accounting Program Year` >= 2012,
                `Accounting Program Description` %in% 
                  c(
                    "LIVESTOCK FORAGE PROGRAM",
                    "LIVESTOCK FORAGE DISASTER PROGRAM"
                  )) |>
  dplyr::group_by(`Program Year` = `Accounting Program Year`, `FSA Code`) |>
  dplyr::summarise(
    `Disbursement Amount` = sum(`Disbursement Amount`, na.rm = TRUE)
  ) |>
  dplyr::collect() |>
  dplyr::mutate(`FSA Code` = as.character(`FSA Code`)) |>
  dplyr::rename(`LFP Payments` = `Disbursement Amount`)

```


### Official LFP Eligibility Records

```{r}
fsa_lfp_eligibility <-
  readr::read_csv(
    "https://data.sustainable-fsa.com/fsa-lfp-eligibility/fsa-lfp-eligibility.csv"
  ) |>
  dplyr::filter(`Disaster Type` == "Drought",
                !is.na(`County Name`)) |>
  dplyr::transmute(`Program Year` =  as.integer(`Program Year`), 
                   County = paste0(State, County), 
                   `State Name`, 
                   `County Name`, 
                   `Pasture Type`,
                   `Drought Factor` = as.integer(`Drought Factor`),
                   `Maximum Eligible Payment Months` = as.integer(`Maximum Eligible Payment Months`),
                   `Payment Factor` = as.integer(`Payment Factor`)) |>
  dplyr::arrange(County, `Program Year`, `Pasture Type`, dplyr::desc(`Payment Factor`)) |>
  dplyr::group_by(County, `Program Year`, `Pasture Type`) |>
  dplyr::filter(n() > 1) %>%
  dplyr::arrange(County, `Program Year`, `Pasture Type`, dplyr::desc(`Payment Factor`))
dplyr::filter(`Payment Factor` == max(`Payment Factor`))

```

```{r}
#| label: calc-lfp-eligibility

calc_lfp_eligibility <-
  function(counties){
    usdm_counties_rle <-
      counties %>%
      # Implement a version of run-length encoding
      lazy_dt() %>%
      dplyr::group_by(FIPS) %>%
      # Add a column for end date (`USDM End`) of the week's USDM map (inclusive) and 
      # create an index (`group`) of changes in the county status
      dplyr::mutate(`USDM End` = date + 6,
                    group = cumsum(c(0, diff(usdm_class)) != 0)) %>%
      dplyr::group_by(FIPS, usdm_class, group) %>%
      # By group, calculate the start and end date (inclusive) of the county USDM status, and 
      # the number of weeks in that status.
      dplyr::summarise(
        `USDM Start` = min(date),
        `USDM End` = max(`USDM End`),
        `USDM Weeks` = n(),
        .groups = "drop") %>%
      as_tibble() %>%
      dplyr::mutate(`USDM Interval` = lubridate::interval(start = `USDM Start`,
                                                          end = `USDM End`)) %>%
      dplyr::select(FIPS,  
                    `USDM Start`, `USDM End`, 
                    `USDM Interval`, `USDM Weeks`,
                    USDM = usdm_class) %>%
      dplyr::arrange(FIPS, `USDM Start`)
    
    # First, find runs of USDM categories that overlap with normal grazing periods,
    # and calculate their overlaps. Date intervals in `lubridate` are inclusive of 
    # start and end dates.
    lfp_usdm_calculated <-
      fsa_normal_grazing_period |>
      dplyr::full_join(usdm_counties_rle,
                       by = "FIPS",
                       relationship = "many-to-many") %>%
      # First, filter out rows where the NGP and USDM intervals do not overlap
      # Also, filter periods that are not D2, D3, or D4. They don't matter for LFP.
      dplyr::filter(lubridate::int_overlaps(`Normal Grazing Period`, `USDM Interval`),
                    USDM >= "D2") %>%
      # Then, calculate the intersecting intervals
      dplyr::mutate(
        `LFP Interval` = lubridate::intersect(`Normal Grazing Period`, 
                                              `USDM Interval`)
      ) %>%
      dplyr::select(FIPS, 
                    `Program Year`, `Pasture Type`, `Normal Grazing Period`, 
                    `USDM Interval`, `LFP Interval`, USDM, `USDM Weeks`) %>%
      dplyr::mutate(`LFP Weeks` = (lubridate::time_length(`LFP Interval`, unit = "days") + 1) / 7) %>%
      dplyr::select(FIPS, 
                    `Program Year`, `Pasture Type`, 
                    `LFP Interval`, USDM, `USDM Weeks`, `LFP Weeks`) %>%
      dplyr::arrange(FIPS, `Program Year`, `Pasture Type`) %>%
      dplyr::mutate(
        `Disaster Start Date` = 
          dplyr::case_when(
            USDM == "D2" & `LFP Weeks` >= 8 ~
              as.character(lubridate::int_start(`LFP Interval`) + lubridate::weeks(8)),
            USDM == "D3" ~
              as.character(lubridate::int_start(`LFP Interval`)),
            USDM == "D4" ~
              as.character(lubridate::int_start(`LFP Interval`)),
            .default = as.character(NA)
          ) %>%
          lubridate::as_date()
      )
    
    lfp_payments_calculated <-
      lfp_usdm_calculated %>%
      dplyr::group_by(FIPS, `Program Year`, `Pasture Type`) %>%
      dplyr::filter(!all(is.na(`Disaster Start Date`))) %>%
      dplyr::summarise(
        `Calculated Qualifying Drought Event` = 
          case_when(
            # Rules for program years 2012--2024
            sum(`LFP Weeks`[USDM == "D4"]) >= 4 ~ "D4b",
            sum(`LFP Weeks`[USDM == "D3"]) >= 4 & 
              sum(`LFP Weeks`[USDM == "D4"]) > 0 ~ "D3b & D4a",
            sum(`LFP Weeks`[USDM == "D3"]) >= 4 ~ "D3b",
            sum(`LFP Weeks`[USDM == "D4"]) > 0 ~ "D4a",
            sum(`LFP Weeks`[USDM == "D3"]) > 0 ~ "D3a",
            any(`LFP Weeks`[USDM == "D2"] >= 8) ~ "D2",
            .default = NA
          ),
        `Calculated Disaster Start Date` = 
          min(`Disaster Start Date`, 
              na.rm = TRUE),
        .groups = "drop"
      ) %>%
      dplyr::mutate(
        `Calculated Payments` = 
          dplyr::case_when(
            `Program Year` < 2012 & 
              `Calculated Qualifying Drought Event` %in%
              c("D4b", "D3b & D4a", "D3b", "D4a") ~ 3L,
            `Program Year` < 2012 & 
              `Calculated Qualifying Drought Event` %in%
              c("D3a") ~ 2L,
            `Program Year` < 2012 & 
              `Calculated Qualifying Drought Event` %in%
              c("D2") ~ 1L,
            
            # Calculated payments switched in Program Year 2012
            `Program Year` >= 2012 & 
              `Calculated Qualifying Drought Event` %in%
              c("D4b") ~ 5L,
            `Program Year` >= 2012 & 
              `Calculated Qualifying Drought Event` %in%
              c("D3b & D4a", "D3b", "D4a") ~ 4L,
            `Program Year` >= 2012 & 
              `Calculated Qualifying Drought Event` %in%
              c("D3a") ~ 3L,
            `Program Year` >= 2012 & 
              `Calculated Qualifying Drought Event` %in%
              c("D2") ~ 1L,
            .default = 0L
          )
      )
    
    return(lfp_payments_calculated)
  }
```

## County Definitions

> The Boundary and Annexation Survey (BAS) provides eligible governments, which include tribal, state, and general-purpose local governments, an opportunity to review the Census Bureau’s legal boundary data to confirm the Census Bureau has the correct boundary, name, and status information. BAS also allows participants to review and provide updates to census designated places (CDPs) and linear features. Maintaining correct boundaries helps make sure that the Census Bureau assigns the appropriate housing and population counts to each government. Title 13, Section 6, United States Code authorizes this survey.
>
> The federal government allocates more than \$2.8 trillion in federal funds annually for health, welfare, infrastructure, education, and other federal programs and services. Maintaining correct boundaries helps make sure that the Census Bureau assigns the appropriate housing and population counts to each government so that governments receive funds appropriately and have the best data available for decision-making.
>
> BAS fulfills the agency’s responsibility as part of the National Spatial Data Infrastructure, for which the Office of Management and Budget (OMB) Circular A–16 designates the Census Bureau as the lead federal agency for maintaining national data about legal government boundaries, as well as administrative and statistical boundaries. BAS supports the geospatial data steward responsibilities of the Geospatial Data Act, the Evidence Act, OMB E-Gov, the Federal Geographic Data Committee, Data.gov, GeoPlatform.gov, the National Map, the Geographic Names Information System, and the Geospatial One-Stop.

Boundary and Annexation Survey (BAS) New Incorporations, Mergers, Consolidations, and Disincorporations

[Legal Boundary Change/Annexation Data](https://www.census.gov/geographies/reference-files/time-series/geo/bas/annex.html)


### Legal issues

Codified in [Title 43 U.S. Code Chapter 46](https://uscode.house.gov/view.xhtml?path=/prelim@title43/chapter46&edition=prelim), the Geospatial Data Act (GDA) establishes clear responsibilities for federal agencies:

[§ 2805 – National Geospatial Data Asset Data Themes](https://uscode.house.gov/view.xhtml?req=granuleid:USC-prelim-title43-section2805&num=0&edition=prelim):

- Establishes National Geospatial Data Assets (NGDA) data themes as standardized categories of geospatial data used across federal agencies.
- Designates lead covered agencies responsible for managing each theme's datasets, resources, and services.
- Requires lead agencies to develop standards, coordinate data population efforts, and ensure data quality and accessibility.
- Mandates annual reporting on the status and progress of each NGDA theme.

[§ 2808 – Covered Agency Responsibilities](https://uscode.house.gov/view.xhtml?req=granuleid:USC-prelim-title43-section2808&num=0&edition=prelim):

- Agencies must use geospatial data standards, including metadata standards, and make metadata available through the [GeoPlatform](https://geoplatform.gov).
- Agencies must search existing sources, including the GeoPlatform, before spending federal funds on new data collection.
- Agencies must report annually on their geospatial data activities and compliance. The [2024 USDA Covered Agency Report](https://www.fgdc.gov/gda/gda-ca-reports/fy2024/fy2024-usda-gda-covered-agency-report.pdf) is available on the FGDC website. Notable are the USDA's responses to question 12.1 and 12.2 (pp. 20–21), which acknowledges that the USDA recieves geospatial data from non-Federal sources and asserts that the USDA employs methods to ensure quality in geospatial data collected from non-federal sources.

[§ 2809 – Limitation on Use of Federal Funds](https://uscode.house.gov/view.xhtml?req=granuleid:USC-prelim-title43-section2809&num=0&edition=prelim):

- Federal funds may not be used to collect, produce, acquire, maintain, or disseminate geospatial data unless the data complies with FGDC standards.

In response to the GDA, the Office of Management and Budget released [OMB Circular A–16](https://www.whitehouse.gov/wp-content/uploads/2017/11/Circular-016.pdf), a binding federal policy document that:

- Directs agencies to coordinate geospatial data activities.
- Establishes the Federal Geographic Data Committee (FGDC).
- Defines the National Spatial Data Infrastructure (NSDI) and NGDA themes.

The "Governmental Units and Administrative and Statistical Boundaries" theme is one of the 18 NGDA themes established by OMB Circular A–16 (Appendix E) under the under Section 2805(b)(3)(A) of the Geospatial Data Act.

> **Governmental Units**: These data describe, by a consistent set of rules and semantic definitions, the official boundary of federal, state, local, and tribal governments as reported/certified to the U.S. Census Bureau by responsible officials of each government for purposes of reporting the Nation's official statistics.

It includes core datasets such as:
- County and state boundaries
- Tribal lands
- Congressional districts
- School districts
- Census tracts and blocks

The U.S. Department of Commerce is the Lead Covered Agency (LCA), and the U.S. Census Bureau is the Theme Lead. The current County and Equivalent Entities National TIGER/Line Shapefiles are among the Governmental Units NGDAs listed in the [Governmental Units, and Administrative and Statistical Boundaries Theme data portal](https://ngda-gov-units-geoplatform.hub.arcgis.com). To maintain and improve the Governmental Units NGDAs, the Census Bureau operates the [Boundary and Annexation Survey (BAS)](https://www.census.gov/programs-surveys/bas.html). The BAS website provides its rationale and authorizing legislation (links added):

> The Boundary and Annexation Survey (BAS) provides eligible governments, which include tribal, state, and general-purpose local governments, an opportunity to review the Census Bureau’s legal boundary data to confirm the Census Bureau has the correct boundary, name, and status information. BAS also allows participants to review and provide updates to census designated places (CDPs) and linear features. Maintaining correct boundaries helps make sure that the Census Bureau assigns the appropriate housing and population counts to each government. [Title 13, Section 6, United States Code](https://uscode.house.gov/view.xhtml?req=granuleid:USC-prelim-title13-section6&num=0&edition=prelim) authorizes this survey.
>
> The federal government allocates more than \$2.8 trillion in federal funds annually for health, welfare, infrastructure, education, and other federal programs and services. Maintaining correct boundaries helps make sure that the Census Bureau assigns the appropriate housing and population counts to each government so that governments receive funds appropriately and have the best data available for decision-making.
>
> BAS fulfills the agency’s responsibility as part of the National Spatial Data Infrastructure, for which the Office of Management and Budget (OMB) Circular A–16 designates the Census Bureau as the lead federal agency for maintaining national data about legal government boundaries, as well as administrative and statistical boundaries. BAS supports the geospatial data steward responsibilities of the Geospatial Data Act, the Evidence Act, OMB E-Gov, the Federal Geographic Data Committee, Data.gov, GeoPlatform.gov, the National Map, the Geographic Names Information System, and the Geospatial One-Stop.

Finally, [February 2025 guidance from the National Geospatial Advisory Committee](https://www.fgdc.gov/ngac/meetings/feb-2025/ngac-paper-advancing-the-nations-geospatial.pdf) recommended that priority national datasets, including NGDA themes, be treated as authoritative and shared in ready-to-use formats. It urges FGDC to define "authoritative geospatial data" (p. 6) and monitor its use across agencies.


```{r}
states <- 
  tigris::states(cb = TRUE) %>%
  sf::st_drop_geometry() %>%
  dplyr::select(STATEFP, State = NAME) %>%
  dplyr::arrange(STATEFP)

FIPS <-
  dplyr::full_join(
    readr::read_delim("https://www2.census.gov/geo/docs/reference/codes2020/national_state2020.txt"),
    readr::read_delim("https://www2.census.gov/geo/docs/reference/codes2020/national_county2020.txt")
  )


census_counties <-
  c(
    `2000` = 
      "https://www2.census.gov/geo/tiger/TIGER2010/COUNTY/2000/tl_2010_us_county00.zip",
    `2008` = 
      "https://www2.census.gov/geo/tiger/TIGER2008/tl_2008_us_county.zip",
    `2009` = 
      "https://www2.census.gov/geo/tiger/TIGER2009/tl_2009_us_county.zip",
    `2010` = 
      "https://www2.census.gov/geo/tiger/TIGER2010/COUNTY/2010/tl_2010_us_county10.zip",
    2011:2024 %>%
      magrittr::set_names(.,.) %>%
      purrr::map_chr(
        \(x){
          paste0(
            "https://www2.census.gov/geo/tiger/TIGER",x,"/COUNTY/tl_",x,"_us_county.zip"
          )
        }
      )
  ) %>%
  {
    dplyr::mutate(curl::multi_download(urls = .,
                                       destfiles = 
                                         file.path("data-raw","counties","census", basename(.)),
                                       resume = TRUE),
                  `Census Year` = names(.))
  } %>%
  dplyr::transmute(`Census Year`,
                   Source = url,
                   File = fs::path_rel(destfile)
  ) %>%
  dplyr::arrange(`Census Year`) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    `Release Date` =
      magrittr::set_names(File, File) %>%
      purrr::map_dfr(archive::archive, .id = "directory") %>%
      dplyr::filter(stringr::str_detect(path, ".shp.xml|.shp.iso.xml")) %>%
      # dplyr::arrange(path) %>%
      dplyr::group_by(directory) %>%
      dplyr::summarise(path = dplyr::first(path)) %>%
      dplyr::rowwise() %>%
      dplyr::mutate(
        xml = list(
          archive::archive_read(archive = directory, file = path) %>%
            xml2::read_xml() %>%
            xml2::as_list()
        ),
        date = ifelse(
          stringr::str_detect(path, ".shp.xml"),
          xml$metadata$metainfo$metd[[1]],
          xml$MI_Metadata$dateStamp$Date[[1]]
        ) %>%
          lubridate::as_date()
      ) %>%
      dplyr::mutate(
        `Census Year` = str_extract(directory, "\\d{4}"),
        `Census Year` = ifelse(stringr::str_detect(directory, "county00"),
                               2000, 
                               `Census Year`) %>%
          as.integer(),
        ## No date specified for 2008, so assume May 7 as on 2009--2011
        date = ifelse(`Census Year` == 2008, lubridate::as_date("2008-05-07"), date),
        ## Set effective date of Census 2000 to January 1.
        date = ifelse(`Census Year` == 2000, lubridate::as_date("2000-01-01"), date),
      ) %$%
      date %>%
      lubridate::as_date()
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    Parquet = 
      magrittr::set_names(File, `Census Year`) %>%
      purrr::imap_chr(\(x, year){
        outfile <- 
          file.path("data-derived", "counties", paste0("Census ", year,".parquet"))
        
        if(!file.exists(outfile))
          x %>%
          file.path("/vsizip", .) %>%
          sf::read_sf() %>%
          dplyr::select(
            STATEFP = dplyr::starts_with("STATEFP"),
            COUNTYFP = dplyr::starts_with("COUNTYFP"),
            County = dplyr::starts_with("NAME") & 
              !dplyr::starts_with("NAMELSAD"),
            LSAD = dplyr::starts_with("NAMELSAD")
          ) %>%
          dplyr::left_join(states) %>%
          dplyr::mutate(
            Code = paste0(STATEFP, COUNTYFP),
            County = iconv(County, from = "latin1", to = "UTF-8"),
            State = iconv(State, from = "latin1", to = "UTF-8"),
            LSAD = iconv(LSAD, from = "latin1", to  = "UTF-8")
          ) %>%
          dplyr::select(Code, County, State, LSAD) %>%
          sf::st_cast("MULTIPOLYGON") %>%
          sf::st_cast("POLYGON", warn = FALSE, do_split = TRUE) %>%
          sf::st_make_valid() %T>%
          {suppressMessages(sf::sf_use_s2(FALSE))} %>%
          sf::st_make_valid() %T>%
          {suppressMessages(sf::sf_use_s2(TRUE))} %>%
          # Group by class and generate multipolygons
          dplyr::group_by(Code, County, State, LSAD) %>%
          dplyr::summarise(.groups = "drop",
                           is_coverage = TRUE) %>%
          sf::st_cast("MULTIPOLYGON", warn = FALSE) %>%
          sf::st_transform("EPSG:4326") %>%
          dplyr::mutate(Area = sf::st_area(geometry)) %>%
          dplyr::mutate(`County File` = paste0("Census ", year)) %>%
          dplyr::select(`County File`, Code, County, State, LSAD, Area) %>%
          sf::write_sf(
            outfile,
            driver = "Parquet",
            layer_options = c("COMPRESSION=BROTLI",
                              "GEOMETRY_ENCODING=GEOARROW",
                              "WRITE_COVERING_BBOX=NO"),
          )
        
        return(outfile)
      })
  )







counties <-
  c(
    
    # NDMC Albers
    curl::multi_download(
      urls = 
        "https://data.sustainable-fsa.com/ndmc-counties-albers/Albers.gdb.zip",
      destfiles = file.path("data-raw", "counties", "Albers.gdb.zip"),
      resume = TRUE) %$%
      destfile %>% 
      fs::path_rel() %>%
      sf::read_sf(layer = "counties_detailed_all_2021") %>%
      dplyr::filter(stringr::str_detect(CountyName, "Kootenai")) %>%
      print(n = 400)
    dplyr::select(County_FIPS = CountyFIPS) %>%
      sf::st_transform("EPSG:4326") %>%
      dplyr::arrange(County_FIPS) %>%
      list(`NDMC Albers` = .) %>%
      purrr::imap(
        \(x, y){
          outfile <- file.path("data-derived", "counties", paste0(y, ".parquet"))
          
          if(!file.exists(outfile))
            x %>%
            sf::st_set_geometry("geometry") %>%
            dplyr::mutate(Total = sf::st_area(geometry) %>%
                            units::drop_units()
            ) %>%
            dplyr::select(!geometry) %>%
            sf::write_sf(
              outfile,
              driver = "Parquet",
              layer_options = c("COMPRESSION=BROTLI",
                                "GEOMETRY_ENCODING=GEOARROW",
                                "WRITE_COVERING_BBOX=NO"),
            )
          return(outfile)
        }
      ),
    
    # FSA dd17
    curl::multi_download(
      urls = 
        "https://data.sustainable-fsa.com/fsa-counties-dd17/FSA_Counties_dd17.gdb.zip",
      destfiles = file.path("data-raw", "counties", "FSA_Counties_dd17.gdb.zip"),
      resume = TRUE) %$%
      destfile %>% 
      fs::path_rel() %>%
      sf::read_sf() %>%
      # dplyr::filter(FSA_STCOU != FIPS_C) %>%
      dplyr::filter(FSA_STCOU == "72113") %>%
      dplyr::arrange(FSA_STCOU, FIPS_C) %>%
      print(n = 200)
    dplyr::select(County_FIPS = state_county_fips_code) %>% 
      
      {
        # Round-trip to geojson to get rid of strange geometry
        tmp <- tempfile(fileext = ".geojson")
        sf::write_sf(., tmp,
                     delete_dsn = TRUE)
        sf::read_sf(tmp)
      } %>%
      dplyr::arrange(County_FIPS) %>%
      sf::st_transform("EPSG:4326") %>%
      dplyr::group_by(County_FIPS) %>%
      dplyr::summarise(.groups = "drop") %>%
      sf::st_cast("MULTIPOLYGON") %>%
      sf::st_cast("POLYGON", warn = FALSE, do_split = TRUE) %>%
      sf::st_make_valid() %T>%
      {suppressMessages(sf::sf_use_s2(FALSE))} %>%
      sf::st_make_valid() %T>%
      {suppressMessages(sf::sf_use_s2(TRUE))} %>%
      # Group by class and generate multipolygons
      dplyr::group_by(County_FIPS) %>%
      dplyr::summarise(.groups = "drop",
                       is_coverage = TRUE) %>%
      sf::st_cast("MULTIPOLYGON", warn = FALSE) %>%
      list(`FSA dd17` = .) %>%
      purrr::imap(
        \(x, y){
          outfile <- file.path("data-derived", "counties", paste0(y, ".parquet"))
          
          if(!file.exists(outfile))
            
            x %>%
            sf::st_set_geometry("geometry") %>%
            dplyr::mutate(Total = sf::st_area(geometry) %>%
                            units::drop_units()
            ) %>%
            dplyr::select(!geometry) %>%
            sf::write_sf(
              outfile,
              driver = "Parquet",
              layer_options = c("COMPRESSION=BROTLI",
                                "GEOMETRY_ENCODING=GEOARROW",
                                "WRITE_COVERING_BBOX=NO"),
            )
          
          return(outfile)
        }
      )
  )

```

## US Drought Monitor Archive

```{r}
dir.create(
  file.path("data-raw","usdm"),
  showWarnings = FALSE,
  recursive = TRUE
)

## USDM Archive
jsonlite::fromJSON(
  "https://data.sustainable-fsa.com/usdm/usdm-manifest.json"
) %>%
  dplyr::filter(stringr::str_detect(path, "parquet")) %$%
  path %>%
  file.path("https://data.sustainable-fsa.com", "usdm", .) %>%
  curl::multi_download(urls = .,
                       destfiles = file.path("data-raw", "usdm", basename(.)),
                       resume = TRUE)

```

## Intersection of USDM and Counties

- Use the most authoritative version of county boundaries (US Census)
- Available at the start of the Program Year (start date of earliest normal grazing period)
- Consistent over the duration of all applicable normal grazing periods during the program year


```{r}
## Intersection of Counties and USDM
extract_usdm_counties <-
  function(usdm_sf, counties_sf, outfile){
    
    if(!file.exists(outfile))
      test <- sf::st_intersection(
        counties_sf %>%
          sf::`st_agr<-`("constant"),
        usdm_sf %>%
          sf::`st_agr<-`("constant")
      ) %>%
        # sf::st_cast("MULTIPOLYGON") %>%
        sf::st_make_valid() %>%
        dplyr::mutate(
          `Percent` = units::drop_units(
            sf::st_area(geometry)
          ) / `Total`
        ) %>%
        sf::st_drop_geometry() %>%
        dplyr::select(Date = date, County_FIPS, USDM = usdm_class, Percent) %>%
        dplyr::arrange(Date, County_FIPS, USDM, Percent) %>%
        arrow::write_parquet(sink = outfile,
                             version = "latest",
                             compression = "zstd",
                             use_dictionary = TRUE)
    
    return(outfile)
  }



counties <-
  tibble::tibble(
    County_Path = fs::dir_ls("data-derived/counties",
                             recurse = TRUE,
                             type = "file")
  ) %>%
  dplyr::mutate(County_File = 
                  basename(County_Path) |>
                  stringr::str_remove(".parquet")
  ) 

counties_full <-
  counties %>%
  dplyr::filter(
    County_File %in% 
      c(
        "Census 2000",
        "Census 2010",
        "Census 2020",
        "Census 2024",
        "FSA dd17",
        "NDMC Albers"
      )
  )

usdm <-
  tibble::tibble(
    USDM = fs::dir_ls(file.path("data-raw", "usdm"))
  ) %>%
  dplyr::mutate(Date = 
                  str_extract(USDM, "\\d{4}-\\d{2}-\\d{2}") |>
                  lubridate::as_date()) |>
  dplyr::select(Date, USDM) |>
  dplyr::arrange(Date)


process_counties <-
  function(usdm, counties){
    the.counties <-
      sf::read_sf(counties)
    
    future::plan(future.callr::callr)
    # future::plan(sequential)
    
    usdm %<>%
      dplyr::mutate(
        out = 
          furrr::future_map2_chr(
            USDM, outfile,
            \(USDM, outfile){
              extract_usdm_counties(
                usdm_sf = sf::read_sf(USDM),
                counties_sf = the.counties,
                outfile = outfile
              )
              
            },
            .options = furrr::furrr_options(
              packages = c("sf")
            )
          )
      )
    
    plan(sequential)
    
    return(usdm)
  }

`Census Lagged` <- 
  # Get the effective date of the program year, or the earliest NGP start 
  # date for the year.
  fsa_normal_grazing_period %>%
  dplyr::filter(`Program Year` >= 2012) %>%
  dplyr::group_by(`Program Year`) %>%
  dplyr::summarise(
    `Normal Grazing Period Start Date` = 
      min(`Normal Grazing Period Start Date`),
    `Normal Grazing Period End Date` = 
      max(`Normal Grazing Period End Date`)
  )
# Find the most recent Census counties dataset released prior to the NGP start
census_release_dates <-
  list.files("../usdm-counties/data/census/raw",
             pattern = ".zip",
             full.names = TRUE) %>%
  magrittr::set_names(.,.) %>%
  purrr::map_dfr(archive::archive, .id = "directory") %>%
  dplyr::filter(stringr::str_detect(path, ".shp.xml|.shp.iso.xml")) %>%
  dplyr::arrange(path) %>%
  dplyr::group_by(directory) %>%
  dplyr::summarise(path = dplyr::first(path)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    xml = list(
      archive::archive_read(archive = directory, file = path) %>%
        xml2::read_xml() %>%
        xml2::as_list()
    ),
    date = ifelse(
      stringr::str_detect(path, ".shp.xml"),
      xml$metadata$metainfo$metd[[1]],
      xml$MI_Metadata$dateStamp$Date[[1]]
    ) %>%
      lubridate::as_date()
  ) %>%
  dplyr::mutate(
    `Census Year` = str_extract(directory, "\\d{4}"),
    `Census Year` = ifelse(stringr::str_detect(directory, "county00"),
                           2000, 
                           `Census Year`) %>%
      as.integer()
  )

#.shp.xml
`Census Release Dates`$xml[[6]]$metadata$metainfo$metd[[1]]

#.shp.iso.xml
`Census Release Dates`$xml[[11]]$MI_Metadata$dateStamp$Date[[1]]


archive::archive("../usdm-counties/data/census/raw//tl_2010_us_county.zip")

usdm_counties <- 
  cross_join(usdm, counties_full) %>%
  dplyr::bind_rows(
    dplyr::left_join(
      usdm %>%
        dplyr::mutate(Year = lubridate::year(Date)),
      counties %>%
        dplyr::filter(stringr::str_starts(County_File, "Census")) %>%
        dplyr::mutate(Year = str_extract(County_File, "\\d{4}") %>%
                        as.integer() + 1L) %>%
        
        tidyr::complete(Year = 1999:(lubridate::year(lubridate::today()))) %>%
        tidyr::fill(County_Path, County_File) %>%
        tidyr::fill(County_Path, County_File, .direction = "up")
    ) %>%
      dplyr::mutate(County_File = "Census Lagged") %>%
      dplyr::select(-Year)
  ) %>%
  dplyr::mutate(outfile = 
                  file.path("data-derived", "usdm", 
                            paste0("County_File=",County_File), 
                            paste0("Date=", Date),
                            paste0(County_File, "_", Date, ".parquet")
                  )
  ) %>%
  dplyr::arrange(Date, County_File) %T>%
  {
    .$outfile %>%
      dirname() %>%
      unique() %>%
      purrr::walk(
        \(x){
          dir.create(x,
                     recursive = TRUE,
                     showWarnings = FALSE)
        }
      )
  } %>%
  tidyr::nest(usdm = !c(County_Path)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    usdm_processed = list(process_counties(usdm = usdm, 
                                           counties = County_Path))
  ) %>%
  dplyr::select(usdm_processed) %>%
  tidyr::unnest(usdm_processed) %>%
  dplyr::select(County_File, out) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(out = list(arrow::read_parquet(out))) %>%
  tidyr::unnest(out) %>%
  tidyr::pivot_wider(names_from = County_File,
                     values_from = Percent) %T>%
  arrow::write_parquet(sink = file.path("data-derived", "usdm_counties.parquet"),
                       version = "latest",
                       compression = "zstd",
                       use_dictionary = TRUE)

```

## County Effective USDM Class

```{r}
usdm_counties <-
  arrow::read_parquet(file.path("data-derived", "usdm_counties.parquet")) %>%
  tidyr::pivot_longer(!c(Date, County_FIPS, USDM),
                      values_to = "Percent",
                      names_to = "County_File") %>%
  # Filter classes that don't intersect with county boundaries
  dplyr::filter(Percent != 0) %>%
  # Drop the intersection percent
  dplyr::select(-Percent) %>%
  # Convert to data-table for fast summaries
  lazy_dt() %>%
  # get the largest USDM class in each county/date
  dplyr::summarise(USDM = max(USDM, na.rm = TRUE),
                   .by = !c(USDM)) %>%
  as_tibble() %>%
  # Join with a complete accounting of USDM dates by county by county file
  dplyr::right_join(
    counties_full %>%
      dplyr::rowwise() %>%
      dplyr::mutate(sf = list(
        sf::read_sf(County_Path) %>%
          sf::st_drop_geometry() %>%
          dplyr::select(County_FIPS)
      )) %$%
      magrittr::set_names(sf, County_File) %>%
      dplyr::bind_rows(.id = "County_File") %>%
      dplyr::cross_join(
        usdm_dates %>%
          dplyr::select(Date = `USDM Date`)
      )
  ) %>%
  # Create "None" USDM class for counties not affected by drought or abnormally dry conditions
  dplyr::mutate(USDM = tidyr::replace_na(USDM, "None"),
                # Convert to an ordered factor
                USDM = factor(USDM,
                              levels = c("None", paste0("D", 0:4)),
                              ordered = TRUE)) %>%
  dplyr::select(County_File, County_FIPS, Date, USDM) %>%
  dplyr::arrange(County_File, County_FIPS, Date)


usdm_counties_rle <-
  usdm_counties %>%
  # Implement a version of run-length encoding
  lazy_dt() %>%
  dplyr::group_by(County_File, County_FIPS) %>%
  # Add a column for end date (`USDM End`) of the week's USDM map (inclusive) and 
  # create an index (`group`) of changes in the county status
  dplyr::mutate(`USDM End` = Date + 6,
                group = cumsum(c(0, diff(USDM)) != 0)) %>%
  dplyr::group_by(County_File, County_FIPS, USDM, group) %>%
  # By group, calculate the start and end date (inclusive) of the county USDM status, and 
  # the number of weeks in that status.
  dplyr::summarise(
    `USDM Start` = min(Date),
    `USDM End` = max(`USDM End`),
    `USDM Weeks` = n(),
    .groups = "drop") %>%
  as_tibble() %>%
  dplyr::mutate(`USDM Interval` = lubridate::interval(start = `USDM Start`,
                                                      end = `USDM End`)) %>%
  dplyr::select(County_File,
                County = County_FIPS,  
                `USDM Start`, `USDM End`, 
                `USDM Interval`, `USDM Weeks`,
                USDM) %>%
  dplyr::arrange(County_File, County, `USDM Start`)

```

## LFP Payment Calculation

1-LFP Rev 6, Paragraph 23B says:

> The National Office receives drought data from the U.S. Drought Monitor on a weekly basis. The determination of the beginning date of a qualifying drought by pasture type is based on the start and ending dates of a D2, D3, or D4 drought intensity ratings on the U.S. Drought Monitor weekly report data. The U.S. Drought Monitor report is created weekly on Tuesdays, released on Thursdays, and is effective until the following Monday. For D2, D3b, and D4b qualifying drought, the start date of qualifying drought, for counting the duration of consecutive or non-consecutive weeks for eligibility purposes, is the later of either the Tuesday report date or the beginning date of the grazing period if it falls within that week's report. The end date is the required number of weeks from the start date for eligibility purposes, which is the date of qualifying drought. Since the duration requirement for D3a and D4a is 1 day, the date of qualifying drought is the later of either the beginning date of the week’s qualifying report, or the first date of the grazing season, if it falls within that qualifying week’s report. The date of qualifying drought as determined, must fall during the established grazing period for the county and pasture type.

So, according to FSA guidance, weeks of eligibility begin on the start date of a qualifying drought condition within the normal grazing period. In other words, a consecutive eight week run of D2, or four weeks of D3 or D4 (not necessarily consecutive), must occur within the normal grazing period, and therefore the intersection of USDM drought class intervals and the normal grazing period intervals is critical.

The beginning date of the qualifying drought is important too. 1-LFP Rev 6, Paragraph 23B continues:

> The beginning date of a qualifying drought is used to determine livestock eligibility, according to paragraph 25, and grazing land eligibility according to paragraph 26. The beginning date of a qualifying drought for a specific type of grazing land or pastureland is the earliest date of the following:
>
> -   the **last day** of the 8-consecutive-week period during the normal grazing period for the specific type of pasture that the county was rated D2 on the U.S. Drought Monitor
> -   the **first day** that a county was rated a D3 or D4 on the U.S. Drought Monitor during the normal grazing period for the specific type of grazing land.

```{r}
# First, find runs of USDM categories that overlap with normal grazing periods,
# and calculate their overlaps. Date intervals in `lubridate` are inclusive of 
# start and end dates.
lfp_usdm_calculated <-
  fsa_normal_grazing_period |>
  dplyr::full_join(usdm_counties_rle,
                   by = "County",
                   relationship = "many-to-many") %>%
  # First, filter out rows where the NGP and USDM intervals do not overlap
  # Also, filter periods that are not D2, D3, or D4. They don't matter for LFP.
  dplyr::filter(lubridate::int_overlaps(`Normal Grazing Period`, `USDM Interval`),
                USDM >= "D2") %>%
  # Then, calculate the intersecting intervals
  dplyr::mutate(
    `LFP Interval` = lubridate::intersect(`Normal Grazing Period`, 
                                          `USDM Interval`)
  ) %>%
  dplyr::select(County_File, County, 
                `Program Year`, `Pasture Type`, `Normal Grazing Period`, 
                `USDM Interval`, `LFP Interval`, USDM, `USDM Weeks`) %>%
  dplyr::mutate(`LFP Weeks` = (lubridate::time_length(`LFP Interval`, unit = "days") + 1) / 7) %>%
  dplyr::select(County_File, County, 
                `Program Year`, `Pasture Type`, 
                `LFP Interval`, USDM, `USDM Weeks`, `LFP Weeks`) %>%
  dplyr::arrange(County_File, County, `Program Year`, `Pasture Type`) %>%
  dplyr::mutate(
    `Disaster Start Date` = 
      dplyr::case_when(
        USDM == "D2" & `LFP Weeks` >= 8 ~
          as.character(lubridate::int_start(`LFP Interval`) + lubridate::weeks(8)),
        USDM == "D3" ~
          as.character(lubridate::int_start(`LFP Interval`)),
        USDM == "D4" ~
          as.character(lubridate::int_start(`LFP Interval`)),
        .default = as.character(NA)
      ) %>%
      lubridate::as_date()
  )

lfp_payments_calculated <-
  lfp_usdm_calculated %>%
  dplyr::group_by(County_File, County, `Program Year`, `Pasture Type`) %>%
  dplyr::filter(!all(is.na(`Disaster Start Date`))) %>%
  dplyr::summarise(
    `Calculated Qualifying Drought Event` = 
      case_when(
        # Rules for program years 2012--2025
        sum(`LFP Weeks`[USDM == "D4"]) >= 4 ~ "D4b",
        sum(`LFP Weeks`[USDM == "D3"]) >= 4 & 
          sum(`LFP Weeks`[USDM == "D4"]) > 0 ~ "D3b/D4a",
        sum(`LFP Weeks`[USDM == "D3"]) >= 4 ~ "D3b",
        sum(`LFP Weeks`[USDM == "D4"]) > 0 ~ "D4a",
        sum(`LFP Weeks`[USDM == "D3"]) > 0 ~ "D3a",
        any(`LFP Weeks`[USDM == "D2"] >= 8) ~ "D2",
        .default = NA
      ),
    `Calculated Disaster Start Date` = 
      min(`Disaster Start Date`, 
          na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::mutate(
    `Calculated Payments` = 
      dplyr::case_when(
        `Program Year` < 2012 & 
          `Calculated Qualifying Drought Event` %in%
          c("D4b", "D3b/D4a", "D3b", "D4a") ~ 3L,
        `Program Year` < 2012 & 
          `Calculated Qualifying Drought Event` %in%
          c("D3a") ~ 2L,
        `Program Year` < 2012 & 
          `Calculated Qualifying Drought Event` %in%
          c("D2") ~ 1L,
        
        # Calculated payments switched in Program Year 2012
        `Program Year` >= 2012 & 
          `Calculated Qualifying Drought Event` %in%
          c("D4b") ~ 5L,
        `Program Year` >= 2012 & 
          `Calculated Qualifying Drought Event` %in%
          c("D3b/D4a", "D3b", "D4a") ~ 4L,
        `Program Year` >= 2012 & 
          `Calculated Qualifying Drought Event` %in%
          c("D3a") ~ 3L,
        `Program Year` >= 2012 & 
          `Calculated Qualifying Drought Event` %in%
          c("D2") ~ 1L,
        .default = 0L
      )
  )

lfp_payments_comparison <-
  lfp_payments_calculated %>%
  dplyr::arrange(`Program Year`, County, `Pasture Type`) |>
  dplyr::full_join(
    fsa_lfp_eligibility %>%
      dplyr::ungroup() %>%
      dplyr::select(County,
                    `Program Year`,
                    `Pasture Type`,
                    `Official Disaster Start Date` = `Disaster Start Date`,
                    `Official Qualifying Drought Event` = `Qualifying Drought Event`,
                    `Official Payments`) %>% 
      dplyr::arrange(`Program Year`, County, `Pasture Type`, dplyr::desc(`Official Payments`)) %>%
      dplyr::distinct(`Program Year`, County, `Pasture Type`, .keep_all = TRUE)
  )



```

## Examples

### 2024, Phillips County, Arkansas, Short Season Small Grains

```{r}
## Reanalysis shows 5 month payment (D4b)
## official records show 4 month payment (D4a)
lfp_payments_comparison %>%
  dplyr::filter(County == "05107",
                `Program Year` == 2024,
                `Pasture Type` == "Short Season Small Grains"
  )

## County hits D4 for five weeks, starting Dec 26, 2023
lfp_usdm_calculated %>%
  dplyr::filter(County == "05107",
                `Program Year` == 2024,
                `Pasture Type` == "Short Season Small Grains"
  )

## Length of the grazing period is less than 5 months (135 days, ~4.45 months)
## Capped at 4 months? Where is this rule documented?
fsa_normal_grazing_period %>%
  dplyr::filter(County == "05107",
                `Program Year` == 2024,
                `Pasture Type` == "Short Season Small Grains"
  ) %$%
  `Normal Grazing Period` %>%
  lubridate::time_length(unit = "days")

## So, how often does this particular thing happen?
## Calculate months using NGP data and see.
fsa_normal_grazing_period %>%
  dplyr::mutate(
    `NGP Length` = 
      lubridate::time_length(`Normal Grazing Period`,
                             unit = "months")
  ) %>%
  dplyr::select(`Program Year`, County, `Pasture Type`, `NGP Length`) %>%
  dplyr::right_join(lfp_payments_comparison) %>%
  dplyr::select(County_File, County, `Program Year`, `Pasture Type`,
                `Calculated Payments`, `Official Payments`,
                `NGP Length`, `Calculated Qualifying Drought Event`,
                `Official Qualifying Drought Event`) %>%
  dplyr::filter(`Calculated Payments` > `Official Payments`)

```
